Predictive modeling with cross validation

### 2017-08-15 model fitting/cross_validation.Rmd
predicted on test data

```{r, echo = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
```

@@

@@

# backwards step-wise selection with selectCox function
model1 <- pec::selectCox(Surv(time_until_answer, answered) ~ category + n_images + n_tags + new_user + weekday + ampm + text_length + device_length + text_contain_punct + title_questionmark + title_beginwh + capital_text + text_all_lower + prior_effort + gratitude + greeting + update + newline_ratio + max_tagwords + frequent_tag + contain_answered + contain_unanswered, data = train1, rule = "aic")
cox
#results
results1 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + weekday + ampm + device_length + title_questionmark + capital_text + update + contain_answered + contain_unanswered, data = train1)
AIC(results1, k = 2) #65059
summary(results1)

# backward/forward step-wise selection with stepAIC function from MASS package
model2 <- stepAIC(model, direction = "both", k = 2)
# results
results2 <- coxph(Surv(time_until_answer, answered) ~ category + n_images + as.factor(new_user) + weekday + ampm + device_length + title_questionmark + title_beginwh + capital_text + update + newline_ratio + frequent_tag + contain_answered + contain_unanswered, data = train1)
AIC(results2, k = 2) # 65050.45
summary(results2)

@@

# if there are less than 60 questions in a category, group them with other
counts <- train1 %>%
  group_by(new_category) %>%
  summarise(n = n())
for (i in which(counts$n <= 100)) {
  train1$new_category[train1$new_category == counts$new_category[i]] <- "Other"
}

table(train1$new_category)

# model with new_category 
r2_new <- coxph(Surv(time_until_answer, answered) ~ new_category + n_images + new_user + weekday + ampm + device_length + title_questionmark + title_beginwh + capital_text + update + newline_ratio + frequent_tag + contain_answered + contain_unanswered + title_questionmark*title_beginwh + new_user*update + new_user*newline_ratio, data = train1)
AIC(r2_new, k = 2) # 64728.77
summary(r2_new)
```

>>>>

```{r, echo = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(ggplot2)
```

@@

##### Trying out different variable selection functions

@@

# backwards step-wise selection with selectCox function
model1 <- pec::selectCox(Surv(time_until_answer, answered) ~ category + n_images + n_tags + new_user + weekday + ampm + text_length + device_length + text_contain_punct + title_questionmark + title_beginwh + capital_text + text_all_lower + prior_effort + gratitude + greeting + update + newline_ratio + max_tagwords + frequent_tag + contain_answered + contain_unanswered, data = train1, rule = "aic")

#results
results1 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + weekday + ampm + device_length + title_questionmark + capital_text + update + contain_answered + contain_unanswered, data = train1)
AIC(results1, k = 2) #65059
summary(results1)

# backward/forward step-wise selection with stepAIC function from MASS package
model2 <- MASS::stepAIC(first, direction = "both", k = 2)
# results
results2 <- coxph(Surv(time_until_answer, answered) ~ category + n_images + as.factor(new_user) + weekday + ampm + device_length + title_questionmark + title_beginwh + capital_text + update + newline_ratio + frequent_tag + contain_answered + contain_unanswered, data = train1)
AIC(results2, k = 2) # 65050.45
summary(results2)

@@

# if there are less than 60 questions in a category, group them with other
counts <- train1 %>%
  group_by(new_category) %>%
  summarise(n = n())
for (i in which(counts$n <= 100)) {
  train1$new_category[train1$new_category == counts$new_category[i]] <- "Other"
}
table(train1$new_category)
#===================================================

# model with new_category 
final <- coxph(Surv(time_until_answer, answered) ~ new_category + n_images + new_user + weekday + ampm + device_length + title_questionmark + title_beginwh + capital_text + update + newline_ratio + frequent_tag + contain_answered + contain_unanswered + title_questionmark*title_beginwh + new_user*update + new_user*newline_ratio, data = train1)
AIC(final, k = 2) # 64728.77
summary(final)
```

### residuals, PH assumption

```{r}
#=============================================
#martingale residuals
train1$mart <- residuals(final, type = "martingale")
plot_mart <- function(var) {
  ggplot(train1, aes(x = train1[[var]], y = mart)) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Martingale Residuals for", var)) + 
    scale_x_continuous(var)
}

q_vars <- c("n_images", "device_length", "newline_ratio")
purrr::map(q_vars, plot_mart)

#=============================================
#deviance residuals 
train1$deviance <- residuals(final, type = "deviance")
ggplot(train1, aes(x = 1:nrow(train1), y = deviance)) + 
  geom_point() + 
  scale_x_continuous("Question index") + 
  scale_y_continuous(breaks = seq(-4, 4, by = 0.5))
# quite a few questions have a deviance residual of larger than 2.5

#=============================================
#schoenfeld residuals
schoen <- as.data.frame(residuals(final, type = "schoenfeld"))
# complete event times sorted from shortest to longest
schoen$comp_times <- sort(train1[train1$answered != 0,]$time_until_answer)

plot_schoen <- function(var) {
  ggplot(schoen, aes(x = comp_times, y = schoen[[var]])) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Schoenfeld Residuals for", var)) + 
    scale_y_continuous("Residuals")
}
names <- names(schoen)[-31]
purrr::map(names, plot_schoen)

#=============================================
# score residuals
score <- as.data.frame(residuals(final, type = "score"))
# function to plot score residuals
plot_score <- function(column) {
  ggplot(score, aes(x = 1:nrow(score), y = score[[column]])) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Score residuals for", column)) + 
    scale_y_continuous("Score Residuals") + 
    scale_x_continuous("Question index")
}
columns <- names(score)
purrr::map(columns, plot_score)
```

### Predicting on the test data

```{r}
test1 <- x[splitPlan[[1]]$app, ]

#setting up variables in test1
#=============================================
test1$category <- as.character(test1$category)
test1$category[is.na(test1$category)] <- "Other"
#new_category
test1$new_category <- test1$category
apple_terms <- c("apple", "ipod", "ipad")
test1$apple <- str_detect(str_to_lower(test1$device), pattern = START %R% or1(apple_terms) %R% SPC)
test1$new_category[test1$apple == TRUE | test1$subcategory == "iPhone" | test1$category == "Mac"] <- "Apple Product"
test1$new_category[test1$new_category == "Phone"] <- "Android/Other Phone"
test1$new_category[test1$new_category == "Appliance" | test1$new_category == "Household"] <- "Home"
test1$new_category[test1$new_category == "Car and Truck" | test1$new_category == "Vehicle"] <- "Vehicle"
test1$new_category[test1$new_category == "Computer Hardware" | test1$new_category == "Media Player" | test1$new_category == "Apparel"] <- "Other"
# counts <- test1 %>%
#   group_by(new_category) %>%
#   summarise(n = n(), prop = n/nrow(test1))
# for (i in which(counts$prop <= 0.01)) {
#   test1$new_category[test1$new_category == counts$new_category[i]] <- "Other"
# }
#=============================================
#weekday
test1$datetime <- as.POSIXct(test1$post_date,origin="1970-01-01")
test1$weekday <- factor(weekdays(test1$datetime), levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
#=============================================
#ampm
test1$hour <- as.numeric(format(test1$datetime,"%H"))
test1$ampm <- "Night"
test1$ampm[test1$hour >= 5 & test1$hour < 12] <- "Morning"
test1$ampm[test1$hour >= 12 & test1$hour < 17] <- "Afternoon" #noon - 5pm
test1$ampm[test1$hour >= 17 & test1$hour < 20] <- "Evening" #5pm - 8pm
#=============================================
# text length
test1$text_length <- str_length(test1$text)
#=============================================
# device name length
test1$device_length <- str_length(test1$device)
#=============================================
# if the title ends with a question mark
library(rebus)
test1$title_questionmark <- str_detect(test1$title, pattern = QUESTION %R% END)
#=============================================
# if title begins with "wh"
test1$title_beginwh <- str_detect(str_to_lower(test1$title), pattern = "^wh")
#=============================================
test1$capital_text <- str_detect(as.character(test1$text), pattern = "^[[:upper:]]")
#=============================================
test1$update <- str_detect(test1$text, pattern = "===")
#=============================================
test1$newline_ratio <- str_count(test1$text, pattern = "\n")/str_length(test1$text)
#=============================================
# frequent tags
split_tags <- str_split(test1$tags, ", ", simplify = TRUE)
tag_vector <- as.vector(split_tags)
tag_vector <- tag_vector[which(tag_vector != "")]
unique_tags <- unique(tag_vector)

#creating average frequency score variable
test1$tag1 <- split_tags[,1]
test1$tag2 <- split_tags[,2]
test1$tag3 <- split_tags[,3]
test1$tag4 <- split_tags[,4]

assign_score <- function(data, variable) {
  score <- rep(0, nrow(data))
  notempty <- which(data[[variable]] != "")
  for (i in notempty) {
    score[i] <- tag_freq$percent[which(tag_freq$tag == data[[variable]][i])]
  }
  return(score)
}
test1$score1 <- assign_score(test1, "tag1")
test1$score2 <- assign_score(test1, "tag2")
test1$score3 <- assign_score(test1, "tag3")
test1$score4 <- assign_score(test1, "tag4")
test1$avg_tag_score <- (test1$score1 + test1$score2 + test1$score3 + test1$score4)/as.numeric(test1$n_tags)
test1$avg_tag_score[is.nan(test1$avg_tag_score)] <- 0
#=============================================
# if question contains "frequent" tag
percentile80 <- quantile(test1$avg_tag_score, probs = 0.80)

test1$frequent_tag <- FALSE
test1$frequent_tag[test1$avg_tag_score >= percentile80] <- TRUE
#=============================================
#frequent terms in unanswered/answered questions
answered <- test1 %>%
  tbl_df() %>% 
  filter(answered == 1)
unanswered <- test1 %>%
  tbl_df() %>% 
  filter(answered == 0)

test1$contain_unanswered <- str_detect(as.character(test1$title), pattern = or1(freq_terms_u$word))
test1$contain_answered <- str_detect(as.character(test1$title), pattern = or1(freq_terms_a$word))

test1$new_user <- as.factor(test1$new_user)

#356 NAs in the predictions
test_predict <- predict(final, newdata = test1, type = "risk", se.fit = TRUE)
test1$predictions <- test_predict[[1]]

metric <- coxph(Surv(time_until_answer, answered) ~ predictions, data = test1)
summary(metric)
AIC(metric, k = 2)
```
